---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%",
  eval = TRUE
)

```

# geoperu

<!-- badges: start -->

[![CRAN status](https://www.r-pkg.org/badges/version/geoperu)](https://CRAN.R-project.org/package=geoperu)
[![](https://cranlogs.r-pkg.org/badges/grand-total/geoperu?color=green)](https://cran.r-project.org/package=geoperu)
[![](https://cranlogs.r-pkg.org/badges/last-week/geoperu?color=green)](https://cran.r-project.org/package=geoperu)
[![Codecov test coverage](https://codecov.io/gh/PaulESantos/geoperu/branch/main/graph/badge.svg)](https://app.codecov.io/gh/PaulESantos/geoperu?branch=main)
[![R-CMD-check](https://github.com/PaulESantos/geoperu/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/PaulESantos/geoperu/actions/workflows/R-CMD-check.yaml)
[![doi](https://img.shields.io/badge/doi-10.32614/CRAN.package.geoperu-blue.svg)](https://doi.org/10.32614/CRAN.package.geoperu)

[![Lifecycle: stable](https://img.shields.io/badge/lifecycle-stable-brightgreen.svg)](https://lifecycle.r-lib.org/articles/stages.html#stable)
<!-- badges: end -->

The `geoperu` package provides a convenient interface to access official spatial datasets of Peru directly from R. These data are collected from the National Institute of Statistics and Informatics (INEI) of Peru.

## Installation

Install the released version of `geoperu` from CRAN:

``` r
install.packages("geoperu")
```

Install the development version from GitHub with `pak`:

``` r
pak::pak("PaulESantos/geoperu")
```

To access spatial information of districts from any province in Peru, you can use the `get_geo_peru()` function.

```{r example}

library(geoperu)

cusco <- geoperu::get_geo_peru(geography = "CUSCO", 
                               level = "dep",
                               simplified = FALSE,
                               showProgress = FALSE)
cusco
cusco_simplified <- geoperu::get_geo_peru(geography = "CUSCO", 
                                          level = "dep",
                                          simplified = TRUE,
                                          showProgress = FALSE)
cusco_simplified

prov_sf <- geoperu::get_geo_peru(geography = "ANTA",
                                 level = "prov", 
                                 simplified = TRUE,
                                 showProgress = FALSE)
prov_sf
islay_sf <- geoperu::get_geo_peru(geography = "ISLAY",
                                  level = "prov",
                                  simplified = FALSE,
                                  showProgress = FALSE)
islay_sf
```

The objects returned by `geoperu` work directly with `ggplot2`. A restrained
grey scale, fine district borders, and a geographic grid produce a clear,
technical map of Cusco:

```{r}
library(ggplot2)

ggplot() +
  geom_sf(
    data = cusco,
    aes(fill = provincia),
    color = "#FFFFFF",
    linewidth = 0.15
  ) +
  geom_sf(
    data = cusco_simplified,
    fill = NA,
    color = "#374151",
    linewidth = 0.7
  ) +
  scale_fill_grey(start = 0.35, end = 0.8, guide = "none") +
  coord_sf(expand = FALSE) +
  labs(
    title = "CUSCO",
    subtitle = "Provincias y límites distritales",
    caption = "Datos: INEI · Mapa: geoperu"
  ) +
  theme_minimal(base_size = 11) +
  theme(
    plot.background = element_rect(fill = "#FAFAF9", color = NA),
    panel.background = element_rect(fill = "#FAFAF9", color = NA),
    panel.grid = element_line(color = "#E5E7EB", linewidth = 0.3),
    axis.text = element_text(color = "#6B7280", size = 7),
    plot.title = element_text(color = "#1F2937", size = 18, face = "bold"),
    plot.subtitle = element_text(color = "#4B5563"),
    plot.caption = element_text(color = "#6B7280")
  )
```

`geoperu` also provides access to spatial information of natural protected areas in Peru. These areas, managed and declared by the National Service of Natural Areas Protected by the State (SERNANP), encompass a diverse range of ecosystems. The `get_anp_peru()` function allows users to download spatial data representing these protected areas directly into their R environment.


```{r}
manu <- get_anp_peru(anp = "manu", showProgress = FALSE)
manu

manu |> 
  ggplot() +
  geom_sf(fill = "#AAB7B0", color = "#3F5148", linewidth = 0.7) +
  coord_sf(expand = FALSE) +
  labs(
    title = "PARQUE NACIONAL\nDEL MANU",
    subtitle = "Cusco y Madre de Dios",
    caption = "Datos: SERNANP · Mapa: geoperu"
  ) +
  theme_minimal(base_size = 11) +
  theme(
    plot.background = element_rect(fill = "#FAFAF9", color = NA),
    panel.background = element_rect(fill = "#FAFAF9", color = NA),
    panel.grid = element_line(color = "#E5E7EB", linewidth = 0.3),
    axis.text = element_text(color = "#6B7280", size = 7),
    plot.title = element_text(
      color = "#26332D", size = 16, face = "bold", hjust = 0,
      lineheight = 0.9
    ),
    plot.subtitle = element_text(color = "#526159"),
    plot.caption = element_text(color = "#6B7280")
  )

```

These maps use only `ggplot2`; colors, typography, and line widths can be
adapted without modifying the spatial data.
