---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# geoperu

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
[![CRAN status](https://www.r-pkg.org/badges/version/geoperu)](https://CRAN.R-project.org/package=geoperu)
<!-- badges: end -->


The `geoperu` package provides a convenient interface to access official spatial datasets of Peru directly from R. These data are collected from the National Institute of Statistics and Informatics (INEI) of Peru.

## Installation

You can install the development version of `geoperu` as follows:

``` r
pak::pak("PaulESantos/geoperu")
```

To access spatial information of districts from any province in Peru, you can use the `get_geo_peru()` function.

```{r example}

library(geoperu)

cusco <- geoperu::get_geo_peru(geography = "CUSCO", 
                                  level = "dep",
                                  simplified = FALSE)
cusco
cusco_simplified <- geoperu::get_geo_peru(geography = "CUSCO", 
                                  level = "dep",
                                  simplified = TRUE)
cusco_simplified

prov_sf <- geoperu::get_geo_peru(geography = "ANTA",
                                 level = "prov", 
                                 simplified = TRUE)
prov_sf
islay_sf <- geoperu::get_geo_peru(geography = "ISLAY",
                                 level = "prov", 
                                 simplified = FALSE)
islay_sf
```

To visualize the retrieved geospatial information with the `get_geo_peru()` function from the geoperu package, you can use various tools in R. A common option is to use the `ggplot2` library along with `sf` object visualization functionality.

Here's a basic example of how to visualize the data:

```{r}
library(patchwork)
library(ggplot2)
plot1 <- cusco |> 
  ggplot() +
  geom_sf() +
  theme_bw()

plot2 <- cusco_simplified |> 
  ggplot() +
  geom_sf() +
  theme_bw()

plot1 + plot2
```
You can customize the plot by adding additional layers, adjusting styles, and adding labels according to your specific needs.
